Towards a Machine Learning approach for Aggregate Computing

### 2021-10-07 paper-2021-woa-ac-ml.tex
more on RL

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
  twocolumn,
  % hf
]{ceurart}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{xcolor}
\sloppy

@@

\begin{abstract}
Artificial intelligence is here to stay. 
 In particular, Deep Learning techniques in the last years outperform 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
Currently, though, an all-encompassing Deep learning model does not exist, 
 so researchers explore the possibility to apply these techniques in different fields.

@@

\section{Background and Related Work}
\outline{Probably, for a "short" paper, this section can be avoided.}
\outline{walkthrogh traditional engineering way to design CAS}
\outline{walkthrogh Aggregate Computing}

\outline{I do not know, perhaps it is better to avoid this }
\outline{walkthrogh Machine learning approach applied in MARL}
\outline{current limitation (application-specific, not large scale, centralised in some cases,..)}

@@

In this section,
 we analyse the characteristics of the aggregate computing paradigm,
 map these to relevant machine learning contributions in literature,
 pointing out different learning perspective about the integration of Aggregate Computing with Machine Learning,
 and summarise the analysis into a roadmap.

@@

\suggestions{Move the definition of team learning and concurrent learning in the previous section}
\paragraph{Learning problem} As reference to \cite{mas-anc-cooperation}, Multi agent learning can be
 divided in \textit{concurrent} and \textit{team} learning. In the foster, each agent perform the 
 learning process and the in the latter, the learning processo is perform only by one agent an then 
 shared to the collective.
 Team learning is usually used in the ``swarm'' system, due to the homogeneous behaviour -- i.e. each
 agent perform the same program. Here, a common reward signal exist and evaluate the overall collective behaviour.
 Aggregate Computing is quite near to this kind of settings.
 On the other hand, Concurrent Learning, in \cite{csas-and-marl} seems to be usually applied in Collective
 Adaptive System, so again a field near to Aggregate Computing. In particular, Indipendedet Reinforcement Learning
 seems to reach good result in this kind of system, even if no thereom exist as the single agent context.
 Concurrent learning does not presume a global collective behaviour convernce. Indeed using a common reward signal,
 the collective can reach the same behaviour, as in the team learning settings as done here~\cite{iima2008swarm}.
\paragraph{Learning mode}
Aside of what kind of problem we setup, we can use Reinforcement Learning, Supervised Learning.
 Superivsed learning is quite rare in CSAS -- or in the swarm system in general. Novel work leverage Graph Neural network
 to learn how agent should communicate, supposing to know the right behaviour (i.e. produced via global vision simulation).
 In Aggregate Computing, Supervised Learning makes sense if we want to craft new building block by zero. 
 Indeed, we know eventually the right state of the system, but we do not know how the system can reach it

@@

\section{Aggregate Computing with Reinforcement}
In this section, we describe our first result in Aggregate Computing and Machine learning combination,
 in particular, we briefly describe how nodes compute to follow the global specification,
 how Reinforcement Learning can be injected in the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.
\subsection{Computatational Model}
\subsection{Reinforcement Learning integration}
\subsection{Independent Reinforcement Learning}
As a first step, we investigate t
\section{Evaluation} %% optional section
\todo[inline]{define experiment setup, hop count problem}

>>>>

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
  twocolumn,
  % hf
]{ceurart}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{xcolor}
\usepackage{listings}
\sloppy

@@

\begin{abstract}
\suggestions{}
Artificial intelligence is here to stay. 
 In particular, Deep Learning techniques in the last years outperform 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
Currently, though, an all-encompassing Deep learning model does not exist, 
 so researchers explore the possibility to apply these techniques in different fields.

@@

\section{Background and Related Work}
\outline{Probably, for a "short" paper, this section can be avoided.}
\outline{walkthrogh traditional engineering way to design CAS}
\outline{walkthrogh Aggregate Computing}
\subsection{Field calcus}
\subsection{Machine Learning}
\suggestions{I do not know, perhaps it is better to avoid this }
\outline{walkthrogh Machine learning approach applied in MARL}
\outline{current limitation (application-specific, not large scale, centralised in some cases,..)}

@@

In this section,
 we analyse the characteristics of the aggregate computing paradigm,
 map these to relevant machine learning contributions in literature,
 pointing out different learning perspectives about the integration of Aggregate Computing with Machine Learning,
 and summarise the analysis into a roadmap.

@@

\suggestions{Move the definition of team learning and concurrent learning in the previous section}
\paragraph{Learning problem} As reference to \cite{mas-anc-cooperation}, Multi agent learning can be
 divided in \textit{concurrent} and \textit{team} learning. In the foster, each agent perform the 
 learning process and the in the latter, the learning process is performed only by one agent an then 
 shared to the collective.
 Team learning is usually used in the ``swarm'' system, due to the homogeneous behaviour -- i.e. each
 agent perform the same program. Here, a common reward signal exist and evaluate the overall collective behaviour.
 Aggregate Computing is quite near to this kind of settings.
 On the other hand, Concurrent Learning, in \cite{csas-and-marl} seems to be usually applied in Collective
 Adaptive System, so again a field near to Aggregate Computing. In particular, Indipendedet Reinforcement Learning
 seems to reach good result in this kind of system, even if no thereom exist as the single agent context.
 Concurrent learning does not presume a global collective behaviour convernce. Indeed using a common reward signal,
 the collective can reach the same behaviour, as in the team learning settings as done here~\cite{iima2008swarm}.
\paragraph{Learning mode}
Aside of what kind of problem we setup, we can use Reinforcement Learning, Supervised Learning.
 Superivsed learning is quite rare in CSAS -- or in the swarm system in general. Novel work leverage Graph Neural network
 to learn how agent should communicate, supposing to know the right behaviour (i.e. produced via global vision simulation).
 In Aggregate Computing, Supervised Learning makes sense if we want to craft new building block by zero. 
 Indeed, we know eventually the right state of the system, but we do not know how the system can reach it

@@

\section{Aggregate Computing with Reinforcement}
In this section, we describe our first result in Aggregate Computing and Machine learning combination,
 in particular, we briefly describe how nodes compute to follow the global specification,
 how Reinforcement Learning can be injected in the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.

\subsection{Computatational Model}
An aggregate program execution consists in a continuously node-point execution of a \textit{round}, that is
 the atomical part of aggregate computing.
%
Each node can interact only with-in its neighbourhoods sharing the \textit{export}, namely the result produced by
 the evaluation of a round.

The main round steps consist in:
\begin{enumerate}
  \item execution context creation: where node collect information from the sensors, the neighbourhoods exports and the old
  program output
  \item program evaluation: the system-wide aggregate program is evaluated against the local context created by the nodes. This produce 
  the export.
  \item export sharing: the export created will be shared within the current neighbourhoods perceived by the nodes
  \item actuations: giving the export produced, node might be perform some actuation (e.g. motor activations, steering, ...)
\end{enumerate}

The rounds executiong is completly asynchrouns, so do not exist a global clock to coordinate the aggregate. This makes possible
 to scala easly with the node number size.

This collective, proactive and periodical execution will eventually bring to the system-wide specification.

\subsection{Reinforcement Learning integration}
By a local view, a program execution can be see as:
$$
f : CONTEXT => OUTPUT
$$
Where the context contains:
\begin{itemize}
  \item $S(name) => value$: a function from sensor name to the perceived value
  \item $\phi(ID) => EXPORT$ a function from id to neighbourhoods data
\end{itemize}
The computational field $\theta$ contains all output produced by the system.

Speaking by Reinforcement Learning settings, we need to define the state functions
and the reward functions.

Here the execution need to be distributed, so state will be a function from the local node context:
$$
STATE := \Omega(CONTEXT)
$$

Reward function instead, might be depens on the learning mode. Indeed, if we perform learning offline, 
it could be defined from the overall computational field:
$$
REWARD_{global} := \psi({\theta})
$$
If we use a online mode, it need to be defined as a function to the local context:
$$
REWARD_{local} := \psi(\phi)
$$

Finally, actions will be something that influece the current aggregate program output, changing the program evaluation as:
$$
f_{rl} = CONTEXT => \lambda(OUTPUT, ACTION)
$$

As remarked before, the policy $\pi$ need to be distributed so each agent to choose an action perfom:
$$
ACTION = \pi(STATE)
$$

\subsection{Independent Reinforcement Learning}
As a first step, we investigate t
\section{Evaluation} %% optional section
\todo[inline]{define experiment setup, hop count problem}
